---
output:
  pdf_document: default
  html_document: default
---
''[---
title: "Regression Chapter 4 HW"
author: "Pooria Assarehha"
date: "2023-12-22"
output: pdf_document
---

## Ex 4.1

$$\hat{\beta} = (X'X)^{-1} X'Y \quad , \quad X'Y = 
\left[\begin{matrix} 1 & \dots & 1 \\
x_1 & \dots & x_n
\end{matrix} \right]
\left[\begin{matrix}
y_1 \\ \vdots \\ y_n
\end{matrix} \right] = 
\left[\begin{matrix}
\sum_i^n y_i \\ \sum_i^n x_i y_i
\end{matrix} \right]$$

```{r}
xxtinv = matrix(c(28400, -500, -500, 10), nrow=2) / 34000
xty = matrix(c(1100, 61800), nrow=2)
xxtinv %*% xty
```

$$\hat{\beta} = \left[\begin{matrix}
\hat{\beta}_0 \\ \hat{\beta}_1
\end{matrix} \right] = \frac{1}{34000} 
\left[\begin{matrix}
28400 & -500 \\ -500 & 10
\end{matrix} \right] \left[\begin{matrix}
1100 \\ 61800
\end{matrix} \right] =  \left[\begin{matrix}
10 \\ 2
\end{matrix} \right]$$

## Ex 4.4
if $Y$ is a random vector,  define $Cov(Y)$ as variance-covariance matrix

$$Cov(\hat{Y}) = Cov(HY) = HCov(Y)H' = H_{(n\times n)} \sigma^2 I_n H' = \sigma^2 HH'$$
$H$ idempotent and symmetric
$$HH' = \left(X(X'X)^{-1}X'\right)\left(X(X'X)^{-1}X'\right)' = \left(X(X'X)^{-1}X'\right)\left(X \left((X'X)^{-1}\right)' X'\right)  $$

$$ \left(X  \left( (X'X)^{-1}X'X \right)  \left((X'X)^{-1}\right)' X'\right) =   \left(X\left((X'X)'\right)^{-1} X'\right)  = X (X'X)^{-1} X = H'\Rightarrow Cov(\hat{Y}) = \sigma^2 H  $$
$$ \sum_i^n Var(\hat{y}_i) = \sum_i^n \sigma^2 h_{ii} = \sigma^2 tr(H) = \sigma^2tr(X(X'X)^{-1}X') = \sigma^2tr(X'X(X'X)^{-1}) = \sigma^2 tr(I_p) = p\sigma^2$$

## Ex 4.10

```{r}
y = matrix(c(4.7, 4.3, 3.5, 4.0, 3.2, 4.7, 4.3, 4.7, 5.2, 4.2, 3.5, 3.2, 2.6, 
             2.0, 4.0, 3.9, 3.0, 4.5, 2.4), ncol = 1)
x = matrix(c(
  78.0, 67.6, 98.0, 65.6, 65.0, 65.5, 85.5, 76.3, 79.0, 88.2,
  70.5, 75.0, 68.0, 65.0, 70.4, 63.0, 91.2, 67.5, 62.0,
  175, 173, 171, 166, 166, 176, 185, 171, 185, 182, 180, 163, 
  180, 180, 180, 168, 168, 178, 173,
  1, 1, 0, 1, 0, 1, 1, 1, 1, 1, 1, 1, 0, 1, 1, 1, 1, 1, 1, 
  24, 36, 28, 25, 26, 22, 27, 27, 36, 24, 26, 29, 33, 31,
  30, 22, 27, 46, 36), ncol = 4 , byrow =  FALSE)
n = length(y)
p = dim(x)[2]
model = lm(y~x)
summary(model)
```

```{r}
xnew = matrix(c(rep(1,19), x), ncol=5)
B = coef(model)
e = model$residuals
mse = sum(e^2) / (n-p)
for (i in 1:5){
  T = B[i] / sqrt(solve(t(xnew) %*% xnew )[i,i]*mse)
  print(2*pt( abs(T) , df = n-p, lower.tail = FALSE) ) 
}
```

```{r}
R2 = 1 - sum(e^2)/sum( (y - mean(y))^2 )
R2

R2adj = 1 - ((n-1)/(n-p))*(sum(e^2)/sum( (y - mean(y))^2 ))
R2adj
```
Small values of $R^2_{adj}$ means model is not a good fit. even $\bar{y}$ is as good at predicting as this model.

## Ex 4.15
### (a)
$$\hat{\beta} = (X'X)^{-1} X'Y = 
\left[\begin{matrix}
0.1132151 & -0.0044486 & -0.0000836 \\
-0.0044486 & 0.0027437 & -0.0000478 \\ 
-0.0000836 & -0.0000478 & 0.0000012
\end{matrix} \right]
\left[\begin{matrix}
559.60 \\ 7375.44 \\ 337072
\end{matrix} \right] = 
\left[\begin{matrix}
2.365 \\ 1.634 \\ 0.005
\end{matrix} \right] $$

```{r}
xxtinv = matrix(c(0.1132151 , -0.0044486 , -0.0000836 ,
-0.0044486 , 0.0027437 , -0.0000478 , 
-0.0000836 , -0.0000478 , 0.0000012) , nrow = 3)

xty = matrix(c(559.60, 7375.44, 337072), nrow = 3)
B = xxtinv %*% xty
B
```

### (b)

$$ T = \frac{\hat{\beta_2} - \beta_2}{\sqrt{\hat{Var(\hat{\beta}_2)}}} \quad\Rightarrow\quad \frac{1.634}{\sqrt{\frac{SSE}{n-p} \times 0.0027437}}$$

$$SSE = \bf e'\bf e = (\bf y - \bf{Hy} )'(y - Hy) = (y' - y'H')(y - Hy) = y'y - y'Hy$$

$$y'Hy = y' X(X'X)^{-1}X' y = (X'y)' (X'X)^{-1} X'y  $$

```{r}
yty = 18310.6290
sse = yty - t(xty) %*% xxtinv %*% xty
sse
mse = sse / (25 - 3)
mse
```

$$SSE = y'y - y'Hy = 3193.025 \Rightarrow MSE = \frac{3193.025}{25-3} = 145.1375 \quad  \textbf{(d)}$$

```{r}
t = B[2] / sqrt(mse*xxtinv[2,2])
t
```
$$T = \frac{1.634}{\sqrt{145.1375 \times 0.0027437}} = 2.590187$$

### XXXXXX (INCONCLUSIVE) (c)

```{r}
R2adj = 1 - ((24-1)/(25-3))*(sse/(yty - 0 ) )
R2adj
```

$$\bar{y} = \beta_0 + \left[ \begin{matrix} \bar{x}_1 & \bar{x}_2   \end{matrix} \right]\left[ \begin{matrix} \beta_1 \\ \beta_2  \end{matrix} \right]$$
$$R^2_{adj} = 1 - \frac{n-1}{n-p} \frac{SSE}{Syy} = 1 - \frac{25-1}{25-3} \frac{SSE}{y'y - n\bar{y}^2} $$

We need $\bar{y}$ or at least $\bar{X}$ to calculate Syy. MORE INFO NEEDED. PROBLEM IS INCONCLUSIVE. 

### (d)

$$MSE = \frac{SSE}{n-p} = \frac{y'y - y'Hy}{25-3} = 145.1375 $$

